```{r preparation, results="hide",echo=F,message=F,warning=F}

### DEFINED OUTSIDE
# report_dir

library(rmarkdown)
library(knitr)
library(hwriter)
library(magrittr)
options(stringsAsFactors=F)



# Directories and cp data
unlink( file.path(report_dir,"exploratory") , recursive = T)
dir.create( file.path(report_dir,"exploratory"), showWarnings=F,recursive=T)
file.copy(from = "exploratory", to = report_dir, overwrite = T, recursive = T)
zip( file.path(report_dir,"exploratory/all.zip"), files = file.path(report_dir,"exploratory") %>% list.files(full.names=T,recursive=T) )

# Create summary table
summary = file.path(report_dir,"exploratory","index.tsv") %>% read.delim(header=T,check.names=F,colClasses="character",stringsAsFactors=F,comment.char='')
summary = summary[ summary$"File" %>% basename  %in%  # some figures we don't want to show
 c("cordist_hclust_log2CPM.pdf","pca_log2CPM.pdf","top_sd_heatmap_log2CPM.pdf","top_sd_heatmap_cufflinks_logFPKMs.pdf") , ] 
summary = summary[,c("Description","File")]# we want to offer full package as zip
summary %<>% rbind(c("All Figures","exploratory/all.zip"))
summary$"File" %<>% sapply(function(fn){
	hwrite( gsub(".*\\.","",fn) ,link=fn)
})
```

#### Exploratory Analysis 

Different exploratory data analysis (EDA) techniques are applied to gene expression levels, quantified  
by the RSEM log2(CPM) (counts-per-million reads). Main goals of expression level EDA are 
the detection of outliers, potential mislabeling,  to explore the homogeneity of biological replicates and  to appreciate 
the global effects of the different experimental variables. EDA plots are therefore essential to insure clean data is 
sent to differential analysis.

First, we analyze how samples are connected to each other using 
hierarchical clustering on the log2(CPM) gene expression levels. The Ward method is applied 
to the matrix of Pearson's correlation distance.

Secondly, we perform principal component analysis (PCA) on gene expression levels (log2CPM). 
A PCA plot shows the projection of the samples onto a two-dimensional space spanned by the first 
and second principal components. These are the orthogonal directions in which the data exhibits the largest and 
second-largest variability. These two components are usually sufficient to 
differentiate groups of samples describing the principal effects of the 
experimental design.

Next, genes with the most variable expression levels (log2CPM standard deviation across all samples, usually the top 75) 
are used to cluster samples. The utility of such a plot is to get a rough overview of which genes or transcripts are contributing
the most to the samples clustering pattern.

```{r summary_table, results="markup",echo=F,message=F,warning=F}
kable(summary,align=c("l","r"),row.names=F)
```

